Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-02 papers/thesis-kirstie/thesis.tex
edit from last week

The goal of this project is to generate computationally temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
%
%save:
%Outside the region of 
%coexistence which exists between the liquid and crystal states, and which 
%consists of various amounts of liquid and crystal existing together,
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature from 
%which the phase diagrams can be constructed.
%end save
%
%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state. 
%
%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular, 
In general, the Helmholtz free energy, 
which consists of an ideal part and an excess part ,
can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%

Classical Density Functional Theory is used to find the equilibrium 
Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
be used to form the functional for the excess Helmholtz free energy. 
%to predict the freezing of a Weeks-Chandler-Anderson 
%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
%dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
%
%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
According to classical Density Functional Theory the equilbrium 
Helmholtz free energy of the system 
can be found
by varying the number density profile $n(\vec{r})$ until the free energy 
is minimized. 
%The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values of the Helmholtz free energy
%of the system at the given temperature and number density. 
%
%To find the Helmholtz free energy both the ideal Helmholtz free energy
%and the excess Helmholtz free energy must be found and added together.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. 
%
%The excess Helmholtz free energy is that 
%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
%energy $V(r)$ between the spheres.
%
%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%For a spatially varying number density, as would be expected for a crystal, 
%the free energy is a functional of the number density profile $n(\vec{r})$  
%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%analytically to find the resulting free energy for the system. 
%
SFMT makes the task of forming an expression
%is a method used to form an expression 
for the excess Helmholtz free energy of a particular model fluid 
much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
%without 
%having to take explicitly into account the interaction each sphere, or atom, 
%has with every other sphere which would make the task very difficult, 
%or even prohibtive.
%
%Classically, fluids have been modeled by a collection of small, 
%impenetrable spheres colliding with one another like billiard balls. 
%In the simplest case, the spheres are treated as having only a replulsive 
%potential. This model, called the Hard-Sphere fluid, has served as the 
%primary reference system for fluids. This is due to the fact that the 
%repulsion between atoms is much stronger than their attraction, and so 
%the repulsive potential, which is descriptive of hard spheres, dominates 
%the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
%fluid which is a perturbation of the Hard-Sphere fluid created by adding 
%a hard-edge attractive potential to the repulsive potential of the 
%Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
%spheres with softly fading replusive potentials which he modeled by making
%the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
%This reflects the idea that soft-spheres penetrate more deeply into 
%each other when they collide at higher temperatures. 
%
%In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
%with Rowlinson's idea of using temperature-dependent radii to model spheres 
%with soft repulsive potentials to create the Barker-Henderson model 
%fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
%modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
%at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
%Chandler, and Anderson used temperature and density dependent hard-sphere 
%radii to model spheres with the more realistic softly fading repulsive 
%portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
%which is known as the Weeks-Chandler-Anderson model fluid. f
%
%While models for particular fluids were being created, powerful methods 
%to describe the properties of fluids in general were also being developed. 
%In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
%introduced from which an equation of state, and an expression for the 
%excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
%obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
%alternative method...OZ... \color{black}
%%Percus and Yevick took another approach to the Hard-Sphere fluid.
%In 1961 Morita and Hiroike introduced a density functional 
%approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
%(the equilibrium Helmholtz free energy in the case of a canonical 
%ensemble), and the cooresponding equilibrium number density profile for 
%inhomogeneous fluids. To use this approach, a functional must be 
%constructed for a particular fluid.
%%DFT Evans 1979
%%Led by the consistency bewteen results from SPT and Percus-Yevick, 
%In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
%(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
%functional for the Hard-Sphere fluid. Tarazona later improved on the 
%method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
%santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
%forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
%
%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
%with soft repulsive and soft attractive potentials, and developed an erf 
%model for spheres with steep, short-range soft repulsive 
%potentials~\cite{schmidt2000fluid}.
% like that exhibited by a WCA fluid. 
%
Results will be compared to those from Monte-Carlo simulations.

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed~\cite{schroeder}. 
%And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of 
%particles respond to a small offset from equilibrium? 
%
%Save for me:
%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
%\begin{equation}
  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
%\end{equation}
%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
%$dT=0$ gives
%\begin{equation}
  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
%\end{equation}
%where conservation of energy was also applied. Fixing the volume and the 
%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
%results in
%\begin{equation}
  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
%\end{equation}
%which, for fixed N, is the same as
%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
%of particles are fixed, the change in the free energy of the system will 
%always be negative, and thus the free energy tends toward a 
%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
%for a system near equilibrium. 
%End save for me.
A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and that the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft spheres.

@@

\section{Overview}
%
%The goal of this project is to generate phase diagrams for a 
%Weeks-Chandler-Anderson model fluid using a computer program. 
%To do this we find the Helmholtz free energies that the fluid
%would have at a given temperature and density in a  
%liquid state, and in a crystalline solid state. 
%%Once the Helmholtz free energy for both the crystal and liquid states of 
%%the fluid are found for a given temperature and density, they can be 
%%compared to find the resulting state of the fluid. 
%Outside the region of 
%coexistence which exists between the liquid and crystal states, 
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature, 
%and construct phase diagrams.

%%In general, if at some number density and temperature the crystal free energy 
%%is lower than the homogeneous free energy of the liquid, then the fluid will 
%%be in the crystalline state. If the homogeneous liquid free energy is lower 
%%than the crystal free energy, then the fluid will be in the liquid state. 
%
%%To find the equilibrium Helmholtz free energies we use classical Density 
%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%%the corresponding equilibrium number density profile. 
%%In particular, 
%
%In general, the Helmholtz free energy can be written as a functional of 
%the number density profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%Classical Density Functional Theory is used to find the equilbrium 
%Helmholtz free energy of the system at a given temperature and density
%by varying the number density profile $n(\vec{r})$ until the free energy 
%is minimized. The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values for that temperature 
%and number density. 
%
%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
%and the excess Helmholtz free energy.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. The excess Helmholtz free energy is that 
%additional energy which results from the presence of an interatmoic potential 
%energy $V(r)$ between the spheres.
%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%%For a spatially varying number density, as would be expected for a crystal, 
%%the free energy is a functional of the number density profile $n(\vec{r})$  
%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%%analytically to find the resulting free energy for the system. 
%Soft Fundamental Measure Theory is used to form the expression 
%for the excess Helmholtz free energy without having to take explicitly 
%into account the interaction each sphere, or atom, has with every other 
%sphere.

Our goal is to generate computationally temperature-density and pressure-
temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
classical Density Functional Theory. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two discinct regions - one at lower densities, and one at higher densities,
seperated by a coexistence region in which there is found a mixture of the 
fluid at some fixed lower density, and at some fixed higher density. An 
abrupt change from some maximum lower density to some minimum higher density 
occurs for certian combinations of pressure and temperature which form
a line on a pressure-temperature phase diagram. This line distinguishes 
the region in which the fluid exhibits lower densities from the region 
in which the fluid exhbits higher densities. These two regions are known
as the liquid state and the solid state of the fluid.

It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
The transition between the liquid 
and solid states is abruptly inhomogeneous, and we study the behavior 
of the fluid by treating the liquid as homogeneous, and by constructing 
a crystal lattice in real space for the solid. Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site convenietly 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

@@

  %\begin{figure}[h!]
    %\centering
    %% FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
    %%\caption{A Simplified Crystal Cell}
    %\label{fig:simplecell}
  %\end{figure}

%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
%consider instead a simplified crystal cell with one atom in its center 
%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
%minimized for a fixed volume (and temperature and number of particles).  
%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
%be used for illustration. The number density n for a collection of 
%simplified cells is simply 1 atom per fixed bold-box volume.
%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
%It is easier to think of 1 atom per ``box" (a reduced number density) 
%than 1 atom per some length$^3$ (the actual number density). Reduced number 
%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
%of size $\sigma^3$.
%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
%has a volume of $\sigma^3$, the reduced number density is just the number 
%of atoms in the fixed bold-box volume. 
%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
%}
In a large collection of cells, the atom at each lattice site of 
will not realistically reside exactly at the lattice point in the center 
of the cell. Statistically there will typically be a normal distribution 
of positions about the lattice point as pictured in the far right box in 
Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
distribution is shown in Figure \ref{fig:Gaus_plot}. 
A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
a number density profile $n(\vec{r})$  which describes how the atoms are 
statistically distrubuted over a collection of cells. 
%In 3-dimensions, the Gaussian function takes on the form
%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
It is important to note that the number density does not change when introducing a number 
density profile. There are still the same number of atoms over some fixed volume. 
%There is still a total of one atom per ``bold box" volume.
%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping %the volume of the bold box, and thus 
the overall number density the same. This is done by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians of one-eigth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eigth of an 
atom in each small cell. This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
%\begin{align}
  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
%\end{align}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Smallcells}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{SameStatPic.png}
    \caption{Front views of a cube with eight smaller cells. The two views 
    show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
    give rise to the same Gaussian distribution statistically}
    \label{fig:SameStatPic}
  \end{figure}

By varying the width of the Gaussians and the fraction of vacancies, various
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
%There is still a total of one atom per ``bold box" volume. 
%This prodcedure is used only to find the crystal free energy as the number density 
%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
    \caption{Several Gaussian Distribution Number Density Profiles all 
    with the same number density, that is, there is one atom over the fixed volume.}
    \label{fig:Ensemble_vary}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

%The number density profile is varied by varying gw and fv, and the 
For a given combination of number density and temperature, a suite of number 
density profiles are generated.
% like those represented by Figure~\ref{fig:Ensemble_vary}.
The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number density profile,
and also for the homogeneous number denisty profile. 
Both the ideal and excess free energies are found, and added together to get the total Helmholtz 
free energy.
The Helmholtz free energy is minimized for a system with a fixed temperature, 
volume, and number of particles. Setting the number of particles to one, 
the volume can be found from the number density.
The process is then repeated to find the free energies at other combinations 
of number densities and temperatures.

%The number of atoms in one primitive cell is one.

>>>>

Classically, fluids have been modeled by a collection of small, 
impenetrable spheres colliding with one another like billiard balls. 
In the simplest case, the spheres are treated as having only a replulsive 
potential. This model, called the Hard-Sphere fluid, has served as the 
primary reference system for fluids. This is due to the fact that the 
repulsion between atoms is much stronger than their attraction, and so 
the repulsive potential, which is descriptive of hard spheres, dominates 
the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
fluid which is a perturbation of the Hard-Sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading replusive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures.

In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials to create the Barker-Henderson model 
fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods 
to describe the properties of fluids in general were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
alternative method...OZ... \color{black}
%Percus and Yevick took another approach to the Hard-Sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
(the equilibrium Helmholtz free energy in the case of a canonical 
ensemble), and the cooresponding equilibrium number density profile for 
inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency bewteen results from SPT and Percus-Yevick, 
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
functional for the Hard-Sphere fluid. Tarazona later improved on the 
method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft repulsive and soft attractive potentials, and developed an erf 
model for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}. In this paper, Density Functional 
Theory (DFT) will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

The minimization can be illustrated by considering that according to the 
second law of thermodynamics, a system that is not in equilibrium will 
exprience spontanteous processes that increase the total entropy of a 
system and its surroundings until equilibrium is reached, 
$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
the entire universe minus the system. This, together with the conservation 
of energy (the first law of thermodynamics) which says that energy lost 
(or gained) by the system is gained (or lost) by the surroundings 
$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
changes in the free energy will be negative for spontaneous processes 
when the natural variables of the free energy are held fixed. 
%And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of 
%particles respond to a small offset from equilibrium?

Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}
  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
$dT=0$ gives
\begin{equation}
  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
\end{equation}
where conservation of energy was also applied. Fixing the volume and the 
number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
results in
\begin{equation}
  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
\end{equation}
which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
of particles are fixed, the change in the free energy of the system will 
always be negative, and thus the free energy tends toward a 
minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
for a system near equilibrium. A general proof of minimization 
can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

The equations developed so far have been for a hard sphere fluid. All of 
the weight function equations, as well as the equation for the deriviative 
of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
of a delta function characteristic of hard spheres. Soft Fundamental Measure 
Theory (SFMT) extends the Fundamental Measure Theory discussed so far to 
fluids with soft potenials by keeping the relationships between the weight 
functions, and that the weight functions satisfy the same, but replacing 
$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
A new set of weight functions can then be derived for a system of soft spheres.

@@

\section{Overview}
We use classical Density Functional Theory (cDFT) in conjunction 
with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
energy of a system, and the corresponding equilibrium number density profile. 
In particular, SFMT is used to form the expression for the excess free energy 
without having to take into account the interaction each sphere, or atom, 
has with every other sphere. The excess free energy is then added to the 
ideal free energy calculated analytically to find the resulting free energy 
for the system. 
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
For a spatially varying number density, as would be expected for a crystal, 
the free energy is a functional of the number density profile $n(\vec{r})$  
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
the temperature fixed until the free energy is minimized. The minimum free 
energy and its corresponding number density profile are then the equilibrium 
values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
the fluid are found, they can be compared to find the resulting state of 
the fluid. Outside the region of coexistence, the fluid will be completely 
in a crystalline state if the crystal free energy is lower than the liquid 
free energy, and it will be completely in a liquid state if the liquid 
free energy is lower than the crystal free energy.

***** DFT uses n(r) for BOTH Fideal and Fexcess ******

To implement cDFT computationally, we begin by picturing one single crystal 
cell. Although the crystal structure is composed of Face Centered Cubic (FCC) 
cells, the entire lattice can be generated with a cell one-fourth the size 
called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}. 
The primitive cell has the shape of a parallelpiped, and contains the 
equivalent of one atom.

%move this paragraph to methods?
-----------*****-------------

Although the entire crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. The parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, includes portions of eight atoms for a total of one full 
atom, and has eight lattice sites convenietly reachable by moving along 
one of three different lattice vectors to step from one atom to the next. 
The three lattice vectors will be useful in creating the lattice structure 
computationally, and will also make the program run faster.

----------****--------------
%-------------------------------

@@

  \begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
    \caption{A Simplified Crystal Cell}
    \label{fig:simplecell}
  \end{figure}

This cell is rather complicated for the purposes of this discussion, so 
consider instead a simplified crystal cell with one atom in its center 
as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
minimized for a fixed volume (and temperature and number of particles).  
A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
be used for illustration. The number density n for a collection of 
simplified cells is simply 1 atom per fixed bold-box volume.
\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
It is easier to think of 1 atom per ``box" (a reduced number density) 
than 1 atom per some length$^3$ (the actual number density). Reduced number 
density is given by $n^*$ and is equal to the number of atoms in a ``box" 
of size $\sigma^3$.
\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
has a volume of $\sigma^3$, the reduced number density is just the number 
of atoms in the fixed bold-box volume. 
\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}

In a large collection of simple cells, the atom at the center of each cell 
will not realistically reside exactly at the lattice point in the center 
of the cell. Statistically there will typically be a normal distribution 
of positions about the lattice point as pictured in the far right box in 
Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
distribution is shown in Figure \ref{fig:Gaus_plot}. A 3-dimenstional 
Gaussian distribution centered at a lattice point can be used to represent 
a number density profile $n(\vec{r})$  which describes how the atoms are 
statistically distrubuted over a collection of cells.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

@@

It is important to note that the number density does not change when 
introducing a number density profile. There is still a total of one atom 
per ``bold box" volume.

\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}

In a large collection of cells, not all lattice sites will be occupied. 
Vacancies can be taken into account while keeping the volume of the 
bold box, and thus the overall number density, the same. This is done by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians of one-eigth the height.

 \begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Smallcells}
  \end{figure}

Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eigth of an 
atom in each small cell. This makes it convenient to describe the new 
reduced number density in terms of the fraction of vacancies, $f_v$:

\begin{displaymath}{ n^* = (1-f_v){n_{no.vacancies}^*}}\end{displaymath}

\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=2.5cm]{SameStatPic.png}
    \caption{Front views of a cube with eight smaller cells. The two views 
    show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
    give rise to the same Gaussian distribution statistically}
    \label{fig:SameStatPic}
  \end{figure}

By varying the width of the Gaussians and the fraction of vacancies, new 
number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
    \caption{Several Gaussian Distribution Number Density Profiles all 
    with a number density of 1 atom per $V_{bold box}$}
    \label{fig:Ensemble_vary}
  \end{figure}

Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. There is still a total 
of one atom per ``bold box" volume. This prodcedure is used only to find 
the crystal free energy as number density profile for a liquid is 
homogeneous, as shown in Figure \ref{fig:homogen_denisty}.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

A simplified flowchart of the program created to investigate the freezing 
of a WCA fluid is shown below. The program takes four inputs: reduced number 
density n*, reduced temperature t*, a range of Gaussian 
widths gw, and a range of fraction of vacancies fv. The number density 
profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
is calculated for each profile until the minimum free energy is found for 
that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
The process is then repeated to find the free energies at other number 
densities and temperatures.
